---
title: "Assignment 4 - Real Estate"
author: "Bharath Dasari"
date: "21/09/2021"
output: html_document
---

Data Cleansing
```{r}
library(readxl)
REAL_ESTATE_SALES_DATA <- read_excel("REAL ESTATE SALES DATA.xlsx")
View(REAL_ESTATE_SALES_DATA)
real.estate<- as.data.frame(REAL_ESTATE_SALES_DATA)
str(real.estate)
```
```{r}
real.estate$Pool<- as.factor(real.estate$Pool)
real.estate$AC<- as.factor(real.estate$AC)
real.estate$Highway<- as.factor(real.estate$Highway)
real.estate$Quality<-as.factor(real.estate$Quality)
str(real.estate)
View(real.estate)
```

## Cumulative Logit model

```{r}
library(ordinal)
attach(real.estate)
model.fit1<- clm(as.factor(Quality) ~ `Sales price`+`Square feet`+Bedrooms+Bathrooms+AC+
                   Garage+Pool+Year+`lot size`+Highway)
summary(model.fit1)
```

Important Predictors:<br>
1.Sales price<br>
2.Square feet<br>
3.Bedrooms<br>
4.Bathrooms<br>
5.AC<br>
6.Year<br>


Interpretation of Coefficients:<br>

1.	For every unit increase in Sales price the odds of quality of construction being on higher quality changes multiplicatively by a factor of 0.999980340193257 controlling for other variables.<br>
2.	For every unit increase in square feet the odds of quality of construction being on higher quality changes multiplicatively by a factor of 0.998570023392284 controlling for other variables.<br>
3.	For every unit increase in number of Bedrooms the odds of quality of construction being on higher quality changes multiplicatively by a factor of 1.69045884837909 controlling for other variables.<br>
4.	For every unit increase in number of Bathrooms the odds of quality of construction being on higher quality changes multiplicatively by a factor of 0.406000860469356 controlling for other variables.<br>
5.	For every availability of AC, the odds of quality of construction being on higher quality changes multiplicatively by a factor of 0.307586171030328 controlling for other variables.<br>
6.	For every unit increase in the no of cars in Garage the odds of quality of construction being on higher quality changes multiplicatively by a factor of 0.582049373882101 controlling for other variables.<br>
7.	For every availability of Pool, the odds of quality of construction being on higher quality changes multiplicatively by a factor of 1.48839744368959 controlling for other variables.<br>
8.	For every increase in Year the odds of quality of construction being on higher quality changes multiplicatively by a factor of 0.943140514024583 controlling for other variables.<br>







```{r}
ctable <- coef(summary(model.fit1))
p <- round(pnorm(abs(ctable[, "z value"]), lower.tail = FALSE) * 2, digits = 3)
ctable <- cbind(ctable, "p-value" = p)
ci <- confint.default(model.fit1)
ctable <- cbind(ctable, round(exp(cbind(OR = coef(model.fit1), ci)), digits = 3))
ctable
```

1.Sales Price
```{r}
model.fit2<- clm(as.factor(Quality) ~ `Square feet`+Bedrooms+Bathrooms+AC+
                   Garage+Pool+Year+`lot size`+Highway,data=real.estate)
anova(model.fit1,model.fit2)
```

Ho: Sales Price is not a significant variable<br>
Ha: Sales Price is a significant variable<br>

Since p-value is less than α = 0.01, we reject Ho.<br>
Therefore, Sales Price is a  significant variable.<br>

2.Square feet
```{r}
model.fit3<- clm(as.factor(Quality) ~ `Sales price`+Bedrooms+Bathrooms+AC+
                   Garage+Pool+Year+`lot size`+Highway)
anova(model.fit1,model.fit3)
```
Ho: Square feet is not a significant variable<br>
Ha: Square feet is a significant variable<br>

Since p-value is less than α = 0.01, we reject Ho.<br>
Therefore, Square feet is a  significant variable.<br>




3.Bedrooms
```{r}
model.fit4<- clm(as.factor(Quality) ~ `Sales price`+`Square feet`+Bathrooms+AC+
                   Garage+Pool+Year+`lot size`+Highway)
anova(model.fit1,model.fit4)
```
Ho: Bedrooms is not a significant variable<br>
Ha: Bedrooms is a significant variable<br>

Since p-value is less than α = 0.01, we reject Ho.<br>
Therefore, Bedrooms is a  significant variable.<br>



4.Bathrooms
```{r}
model.fit5<- clm(as.factor(Quality) ~ `Sales price`+`Square feet`+Bedrooms+AC+
                   Garage+Pool+Year+`lot size`+Highway)
anova(model.fit1,model.fit5)
```
Ho: Bathrooms is not a significant variable<br>
Ha: Bathrooms is a significant variable<br>

Since p-value is less than α = 0.01, we reject Ho.<br>
Therefore, Bathrooms is a  significant variable.<br>

5.AC
```{r}
model.fit6<- clm(as.factor(Quality) ~ `Sales price`+`Square feet`+Bedrooms+Bathrooms+
                   Garage+Pool+Year+`lot size`+Highway)
anova(model.fit1,model.fit6)
```
Ho: AC is not a significant variable<br>
Ha: AC  is a significant variable<br>

Since p-value is less than α = 0.01, we reject Ho.<br>
Therefore, Bathrooms is a  significant variable.<br>


6.Garage
```{r}
model.fit7<- clm(as.factor(Quality) ~ `Sales price`+`Square feet`+Bedrooms+Bathrooms+AC+
                   Pool+Year+`lot size`+Highway)
anova(model.fit1,model.fit7)
```
Ho: Garage is not a significant variable<br>
Ha: Garage  is a significant variable<br>

Since p-value is greater than α = 0.01, we fail to reject Ho.<br>
Therefore, Garage is a not significant variable.<br>



7.Pool
```{r}
model.fit8<- clm(as.factor(Quality) ~ `Sales price`+`Square feet`+Bedrooms+Bathrooms+AC+
                   Garage+Year+`lot size`+Highway)
anova(model.fit1,model.fit8)
```
Ho: Pool is not a significant variable<br>
Ha: Pool is a significant variable<br>

Since p-value is greater than α = 0.01, we fail to reject Ho.<br>
Therefore, Pool is a not significant variable.<br>


8.Year
```{r}
model.fit9<- clm(as.factor(Quality) ~ `Sales price`+`Square feet`+Bedrooms+Bathrooms+AC+
                   Garage+Pool+`lot size`+Highway)
anova(model.fit1,model.fit9)
```
Ho: Year is not a significant variable<br>
Ha: Year is a significant variable<br>

Since p-value is less than α = 0.01, we reject Ho.<br>
Therefore, Year is a significant variable.<br>

9.lot size
```{r}
model.fit10<- clm(as.factor(Quality) ~ `Sales price`+`Square feet`+Bedrooms+Bathrooms+AC+
                   Garage+Pool+Year+Highway)
anova(model.fit1,model.fit10)
```
Ho: lot size is not a significant variable<br>
Ha: lot size is a significant variable<br>

Since p-value is greater than α = 0.01, we fail to reject Ho.<br>
Therefore, lot size is a not significant variable.<br>

10.Highway
```{r}
model.fit11<- clm(as.factor(Quality) ~ `Sales price`+`Square feet`+Bedrooms+Bathrooms+AC+
                   Garage+Pool+Year+`lot size`)
anova(model.fit1,model.fit11)
```
Ho: Highway is not a significant variable<br>
Ha: Highway is a significant variable<br>

Since p-value is greater than α = 0.01, we fail to reject Ho.<br>
Therefore, Highway is a not significant variable.<br>


Predictors to be retained are: <br>
Important Predictors:<br>
1.Sales price<br>
2.Square feet<br>
3.Bedrooms<br>
4.Bathrooms<br>
5.AC<br>
6.Year<br>


Backward elimination method:

```{r}
library(bootStepAIC)
stepAIC(model.fit1,direction = "backward")
```



```{r}
mod.reduced<-clm(as.factor(Quality) ~ `Sales price` + `Square feet` + Bedrooms + Bathrooms + AC + Garage + Year, data=real.estate)
summary(mod.reduced)
```



```{r}
anova(model.fit1 , mod.reduced , type = 1)

```
Variables to be retained:<br>
Sales price<br>
Square feet<br>
Bedrooms<br>
Bathrooms<br>
AC<br>
Year<br>



## Binary Logistic Regression Model

```{r}

estate.data<- as.data.frame(REAL_ESTATE_SALES_DATA)
estate.data$Pool<- as.factor(estate.data$Pool)
estate.data$AC<- as.factor(estate.data$AC)
estate.data$Highway<- as.factor(estate.data$Highway)
library(car)

estate.data$Quality_High <- recode(estate.data$Quality,"1=1;c(2,3)=0")
estate.data$Quality_Medium <- recode(estate.data$Quality,"2=1;c(1,3)=0")
estate.data$Quality_Low <- recode(estate.data$Quality,"3=1;c(1,2)=0")
str(estate.data)
View(estate.data)

```

```{r}
attach(estate.data)
model.bin1<- glm(Quality_High ~ `Sales price`+`Square feet`+Bedrooms+Bathrooms+AC+
                   Garage+Pool+Year+`lot size`+Highway,family = binomial(link = "logit"))
summary(model.bin1)
```


```{r}
model.bin2<- glm(Quality_Medium ~ `Sales price`+`Square feet`+Bedrooms+Bathrooms+AC+
                   Garage+Pool+Year+`lot size`+Highway,family = binomial(link = "logit"))
summary(model.bin2)
```


```{r}
model.bin3<- glm(Quality_Low ~ `Sales price`+`Square feet`+Bedrooms+Bathrooms+AC+
                   Garage+Pool+Year+`lot size`+Highway,family = binomial(link = "logit"))
summary(model.bin3)
```
Comparison of slope<br>
Variable	        Cumulative	Binary:High Quality 	Binary:Medium Quality	 Binary: Low Quality<br>
Sales price	      -0.000020	    0.00002	           -0.00001	              -0.000013<br>
Square feet	      -0.001431	    0.00093	            0.00052	              -0.002710<br>
Bedrooms	         0.525000	   -0.56660	            0.08308	               0.583400<br>
Bathrooms	        -0.901400	    0.40250	            1.08200	              -1.423000<br>
AC1	              -1.179000	    16.74000	          1.55100	              -1.013000<br>
Garage	          -0.541200	    0.85700	           -0.07346	              -0.422900<br>
Pool1	             0.397700	   -0.38680	            0.06034	               0.941400<br>
Year	            -0.058540	    0.00790	            0.03748	              -0.081500<br>
lot size	        -0.000001	   -0.00002	            0.00004	              -0.000003<br>
Highway1	        -0.433600	    1.45400	           -0.34630	               0.113300<br>



## Final Model

```{r}
mod.final<-clm(as.factor(Quality) ~ `Sales price` + `Square feet` + Bedrooms + Bathrooms + AC + Year + Year, data=real.estate)
summary(mod.final)
```




